---
title: "Project_1"
author: "Yineng Chen"
date: "2020/6/11"
output: 
  html_document:
    keep_md: true
editor_options: 
  chunk_output_type: console
  
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(ggplot2)
library(lattice)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)
```

## Loading and preprocessing the data

```{r}
dir.create("data")
unzip("activity.zip",exdir = "data")
activity = read_csv("./data/activity.csv",col_types = "icc", na = "NA")

activity = activity %>% mutate(date = as.Date(date,  "%Y-%m-%d"))

```

## What is mean total number of steps taken per day?

1. Calculate the total number of steps taken per day

```{r}
step_perday = aggregate(activity$steps, list(activity$date), sum,na.rm = F)

names(step_perday)[1] = "date"
names(step_perday)[2] = "steps"

```

2.  Make a histogram of the total number of steps taken each day

```{r}

ggplot(data = step_perday, aes(x = steps)) + 
  geom_histogram(fill = "pink", binwidth = 1000) +
  labs(title = "Daily Steps", x = "Steps", y = "Frequency")

```

3. Calculate and report the mean and median of the total number of steps taken per day

```{r}
mean_step = mean(step_perday$steps,na.rm = TRUE)
median_step = median(step_perday$steps,na.rm = TRUE)

output1 = cbind(mean_step,median_step)
output1
```


## What is the average daily activity pattern?

Make a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis).

```{r}
step_perinterval = aggregate(activity$steps, list(activity$interval),mean,na.rm = T)

names(step_perinterval)[1] = "intervals"
names(step_perinterval)[2] = "ave_steps"

step_perinterval = step_perinterval %>%  mutate(intervals = as.numeric(intervals))

ggplot(step_perinterval, aes(x = intervals , y = ave_steps, group = 1)) +
  geom_line(color ="blue", size=1) + 
  labs (title = "Avg. Daily Steps", x = "Interval", y = "Average Steps") + theme_bw()
```

2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r}
n = which.max(step_perinterval$ave_steps)
max_interval = step_perinterval$intervals[n]
max_interval
```

## Imputing missing values

1. Calculate and report the total number of missing values in the dataset 
```{r}
sum(is.na(activity$steps))
```

2. Devise a strategy for filling in all of the missing values in the dataset. 
```{r}
activityDT <- data.table::fread(input = "data/activity.csv")

# Filling in missing values with median of dataset. 
activityDT[is.na(steps), "steps"] <- activityDT[, c(lapply(.SD, median, na.rm = TRUE)), .SDcols = c("steps")]
```

3. Create a new dataset that is equal to the original dataset but with the missing data filled in.

```{r}
data.table::fwrite(x = activityDT, file = "data/tidyData.csv", quote = FALSE)
```

4. Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day.

```{r}
tidy_activity = read_csv("data/tidyData.csv", col_types = "icf")%>% 
  mutate(date = as.Date(date,  "%Y-%m-%d"))

step_day_tidy = aggregate(tidy_activity$steps, list(tidy_activity$date), sum,na.rm = T)

names(step_day_tidy)[1] = "date"
names(step_day_tidy)[2] = "steps"

ggplot(data = step_day_tidy, aes(x = steps)) + 
  geom_histogram(fill = "lightgreen", binwidth = 1000) +
  labs(title = "Daily Steps", x = "Steps", y = "Frequency")

# mean and median
mean_step2 = mean(step_day_tidy$steps,na.rm = TRUE)
median_step2 = median(step_day_tidy$steps,na.rm = TRUE)

output2 = cbind(mean_step2,median_step2)
output = rbind(output1,output2)
row.names(output) = c("With NA","Fill in Na")
output
```

## Are there differences in activity patterns between weekdays and weekends?

1. Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.

```{r}
n = nrow(tidy_activity)
Sys.setlocale("LC_TIME", "English")

tidy_activity = tidy_activity %>% mutate(day = weekdays(tidy_activity$date)) %>% mutate(weekdays = vector(mode = "character",length = n))



for(i in 1:n){
  if(tidy_activity$day[i] %in% c("Monday","Tuesday","Wednesday","Thursday", "Friday")){
    tidy_activity$weekdays[i] = "weekday"
  } else {
    tidy_activity$weekdays[i] = "weekend"
  }
}

tidy_activity = tidy_activity %>% mutate(weekdays = as.factor(weekdays))  
```

2. Make a panel plot containing a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). 

```{r results='asis'}

step_interval = aggregate(tidy_activity$steps, list(tidy_activity$interval,tidy_activity$weekdays),mean,na.rm = T)

names(step_interval)[1] = "interval"
names(step_interval)[2] = "weekdays"
names(step_interval)[3] = "ave_steps"

step_interval = step_interval %>% mutate(interval = as.numeric(interval))

p = ggplot(step_interval, 
       aes(x = interval, y = ave_steps, group=1, color = weekdays)) +  
  geom_line() + 
   facet_grid(weekdays~.) +
  labs(x = "Interval",
       y = "Average number of steps", 
       title = "Activity Patterns of Weekdays and Weekends")
p
```

```{r results='hide', include=FALSE}
# Just recreating activityDT from scratch then making the new factor variable. (No need to, just want to be clear on what the entire process is.) 
activityDT <- data.table::fread(input = "data/activity.csv")
activityDT[, date := as.POSIXct(date, format = "%Y-%m-%d")]
activityDT[, `Day of Week`:= weekdays(x = date)]
activityDT[grepl(pattern = "Monday|Tuesday|Wednesday|Thursday|Friday", x = `Day of Week`), "weekday or weekend"] <- "weekday"
activityDT[grepl(pattern = "Saturday|Sunday", x = `Day of Week`), "weekday or weekend"] <- "weekend"
activityDT[, `weekday or weekend` := as.factor(`weekday or weekend`)]
head(activityDT, 10)


activityDT[is.na(steps), "steps"] <- activityDT[, c(lapply(.SD, median, na.rm = TRUE)), .SDcols = c("steps")]
IntervalDT <- activityDT[, c(lapply(.SD, mean, na.rm = TRUE)), .SDcols = c("steps"), by = .(interval, `weekday or weekend`)] 

ggplot(IntervalDT , aes(x = interval , y = steps, color=`weekday or weekend`)) + geom_line() + labs(title = "Avg. Daily Steps by Weektype", x = "Interval", y = "No. of Steps") + facet_wrap(~`weekday or weekend` , ncol = 1, nrow=2)
```

